Stochastic Variance-Reduced Policy Gradient

### 2018-02-09 svrpg.tex

\section{Practical SVRPG versions}\label{app:practicalsvrpg}
In this appendix we provide more details on the practical variants of our algorithm.
\subsection{Learning Rates}
\begin{algorithm}[h]
	\begin{algorithmic}
		\STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
		\STATE $\kappa_t = \beta_1 \kappa_{t-1} + (1 - \beta_1) g_t$
		\STATE $\nu_t = \beta_2 \nu_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
		\STATE $\hat{\kappa}_t = \dfrac{\kappa_t}{1 - \beta^t_1}$
		\STATE $\hat{\nu}_t = \dfrac{\nu_t}{1 - \beta^t_2}$
		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{\nu}_t} + \epsilon} \hat{\kappa}_t$
		\STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
	\end{algorithmic}
	\caption{
		\label{A:adam}
		Adam}
\end{algorithm}
Let us give a deeper insight on the two different learning rate schedules used by our algorithm. We report pseudo-code of the original ADAM in Algorithm \ref{A:adam}. As mentioned, we use two distinct instances of ADAM to manage different sources of variance: one related to the snapshots, and one to the sub-iterations. In this way the ADAM associated to the snapshots takes into account only the history of gradient moments at the snapshots. Using the notation in Algorithm \ref{A:adam}, we can explicitly define our gradient updates:
\begin{align*}
\vtheta^{s+1}_1 &= \wt{\vtheta}^s + \alpha^{\textsc{FG}}_s\left(\wh{\nabla}_N J(\wt{\vtheta}^s) \right)\\
\vtheta^{s+1}_{t+1} &= \vtheta^{s+1}_t + \alpha^{\textsc{SI}}_{s+1,t} \Big( 
%        v_t^{s+1}
\blacktriangledown J(\vtheta^{s+1}_t)\Big)
\text{ for $t=1,\dots,m-1$},
\end{align*}

\subsection{Baseline}
For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
\begin{align*}
\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
\end{align*}
where $s$ is the state vector and $\odot$ is the element-wise product. In the case of our algorithm the model representing the baseline is fitted every snapshot so we have:\\
\begin{align*}
        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
          + \frac{1}{B} \sum_{n=1}^{B}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
\end{align*}
Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration. 
\section{Experimental Details}\label{app:exp}
Used environments:
\begin{enumerate}
	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body. Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$. The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
    \item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles. 2-dimensional action space: the two momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2]{a}^2$. The episodes terminates when the number of time steps T is greater than 500.
    \item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles. 6-dimensional action space: the 6 momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 0.05\norm[2]{a}^2$. The episodes terminates when the number of time steps T is greater than 500.
\end{enumerate}

>>>>

\section{Practical SVRPG versions}\label{app:practicalsvrpg}
In this appendix we provide more details on the practical variants of our algorithm.
\subsection{Learning Rates}
\begin{algorithm}[h]
	\begin{algorithmic}
		\STATE \textbf{Input:} A gradient estimate $g_t$ and parameters $\beta_1$, $\beta_2$, $\epsilon$ and $\alpha$.
		\STATE $m_t = \beta_1 m_{t-1} + (1 - \beta_1) g_t$
		\STATE $v_t = \beta_2 v_{t-1} + (1 - \beta_2) g_t \circ g_t$ ($\circ$ is the Hadamard (component-wise) product)
		\STATE $\hat{m}_t = \dfrac{m_t}{1 - \beta^t_1}$
		\STATE $\hat{v}_t = \dfrac{v_t}{1 - \beta^t_2}$
		\STATE $\Delta(g_t) = \dfrac{\alpha}{\sqrt{\hat{v}_t} + \epsilon} \hat{m}_t$
		\STATE \textbf{Return:} The increment $\Delta(g_t)$  of the parameters.
	\end{algorithmic}
	\caption{
		\label{A:adam}
		Adam}
\end{algorithm}
Let us give a deeper insight on the two different learning rate schedules used by our algorithm. We report pseudo-code of the original ADAM in Algorithm \ref{A:adam}. As mentioned, we use two distinct instances of ADAM: one related to the snapshots, and one to the sub-iterations. In this way the ADAM associated to the snapshots will take into account only the history of gradient moments at the snapshots. The same reasoning holds for the sub-iterations. This, of course, allows our algorithm to take into account the different variabilities in a more accurate manner.
\subsection{Baseline}
For all batch gradient-based algorithms, we use the same time-varying feature encoding for the linear baseline:
\begin{align*}
\psi_{s,t}=concat(s, s \odot s, 0.01t, (0.01t)^2, (0.01t)^3,1)
\end{align*}
where $s$ is the state vector and $\odot$ is the element-wise product. In the case of our algorithm the model representing the baseline is fitted every snapshot so we have:\\
\begin{align*}
        \nabla J(\pi^T) =\frac{1}{N} \sum_{n=1}^{N}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(y_k^n)\right) (\gamma^h \Reward(y_h^n)-b_{sn})
          + \frac{1}{B} \sum_{n=1}^{B}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\\ - \frac{1}{B} \sum_{n=1}^{B}
                \sum_{h=0}^H \left(\sum_{k=0}^h \nabla \log\pi^T(z_k^n)\right) (\gamma^h \Reward(z_h^n) - b_{si})\omega(z_{0:h}|\pi^B,\pi^T)
\end{align*}
Where $b_{sn}$ is the prediction given by the baseline at the snapshot, whereas $b_{si}$ is the prediction given by the baseline at a given subiteration. 
\section{Experimental Details}\label{app:exp}
Used environments:
\begin{enumerate}
	\item \emph{Cart-Pole Balancing} : 4-dimensional state space: cart position x, pole angle $\theta$, cart velocity $\dot{x}$ and pole velocity $\dot{\theta}$; 1-dimensional action space: the horizontal force applied to the cart body. Reward function  is defined as $r(s, a) := 10 - (1 - cos(\theta)) - 10^{-5}\norm[] a^2$. The episodes terminates when $|x|>2.4$ or $|\theta|>0.2$ or the number of time steps T is greater than 100.
    \item \emph{Mujoco Swimmer}: 13-dimensional state space: 3 links velocities ($v_x$ and $v_y$ of center of masses) and 2 actuated joints angles. 2-dimensional action space: the two momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 10^{-4}\norm[2]{a}^2$. The episodes terminates when the number of time steps T is greater than 500.
    \item \emph{Mujoco Half Cheetah}: 20-dimensional state space: 9 links and 6 actuated joints angles. 6-dimensional action space: the 6 momentums applied on actuated joints.  The reward function is defined as $r(s, a) := v_x - 0.05\norm[2]{a}^2$. The episodes terminates when the number of time steps T is greater than 500.
\end{enumerate}
